---
title: "Crime Rates"
---

Follow along with the demo to learn how to calculate crime rates using R and the Tidyverse. We'll also introduce a new library, janitor, which helps make column names easier to work with. You should get started by loading the tidyverse in a new R block, then install and load janitor.

### Task 1: Load the tidyverse and install and load janitor
```{r}
library(tidyverse)
install.packages("janitor")
library(janitor)
```



### Task 2: Load the crime dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_annual_crime.csv

```{r}
Crime_Data <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_annual_crime.csv") |> clean_names()
```


### Task 3: Use janitor to clean up the column names

skip

### Task 4: Calculate the total crime rate for each year (picking our unit!)

```{r}
Crime_Data |> mutate(total_rate = grand_total/population * 10000)
```


### Task 5: Calculate the violent and property crime rates for each year (same unit)

```{r}
Crime_Data |> mutate(violent_rate = violent_crime_total/population * 10000, property_rate = property_crime_total/population * 10000)
```


### Task 6: Describe our results

Write a few sentences about what you found in the data. What years had the highest crime rates? What years had the lowest crime rates? Does the pattern match your expectations?

The first thing that stood out to me is that 2020 had the lowest rates overall in both the violent crime and property crime categories. This makes complete sense, as we were in a global pandemic where people rarely left their houses. The highest crime rates appeared to be in 1995. This also makes sense because police enforcement was not always as relevant as it is today.


When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.

