---
title: "UMD Fees"
---

Follow along with the demo to learn how to calculate percentage change using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.


### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```




### Task 2: Load the UMD Fees dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv

```{r}
umd_fees <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/umd_fees.csv")
```


### Task 3: Calculate the percentage change in each category between Fall 2024 and Fall 2021 using `mutate()`
```{r}
umd_fees_pct <- umd_fees |> mutate(pct_change = (`Fall 2024`-`Fall 2021`)/`Fall 2021` * 100)
```

### Task 4: Describe our results

Write a few sentences about what you found in the data. What's the story here?

We found in the data that the categories with the highest percent increase were student sustainability and shuttle buses. We can assume these fees are being increased simply because the university can. All the categories have been steadily increasing, but these two have seen the most change, even though there is not really any noticeable change through my perspective as a student. Athletics is the only category that has not changed at all, but we can expect this to change once the athletes begin to receive payment. 



When you are done, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
